---
title: "Visualization Part 1"
author: "Devon Morgan"
date: "9/27/2018"
output: html_document
---

Code set up: 

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

# Set global parameters for figures produced today. 

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```

## R Markdown


```{r data_import, cache = TRUE}
library(rnoaa)

weather_df = rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Revisit plot

Always helpful to include labels in plots. This is helpful for x and y axes and overall. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
      title = "Temperature plot",
      x = "Minimum daily temperature (C)",
      y = "Maxiumum daily temperature (C)",
      caption = "Data from the rnoaa package"
    )
```

Next, manipulate scale using `scale` functions. The `scale_x_*` where the value for * depends on the type of variable of interest. Use Option + 0 to make a degree symbol. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
      title = "Temperature plot",
      x = "Minimum daily temperature (C)",
      y = "Maxiumum daily temperature (C)",
      caption = "Data from the rnoaa package"
    ) + 
  scale_x_continuous(
      breaks = c(-15, 0, 15), 
      labels = c("-15ยบ C", "0", "15"),
      limits = c(-20, 30)) +
  # switches the y axis to the other side 
  scale_y_continuous(trans = "sqrt", 
                     position = "right")

```

## Colors and themes

Adjust color. Can rename the name that appears in the legend using `name` function in `scale_color_hue`. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
      title = "Temperature plot",
      x = "Minimum daily temperature (C)",
      y = "Maxiumum daily temperature (C)",
      caption = "Data from the rnoaa package"
    ) + 
  scale_color_hue(name = "Location",
                    h = c(100, 200),
                    l = 50)
```

Use pre-described color pallettes, namely `viridis` package. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
      title = "Temperature plot",
      x = "Minimum daily temperature (C)",
      y = "Maxiumum daily temperature (C)",
      caption = "Data from the rnoaa package"
    ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  )
```

Can also change the position of the plot legend: 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  theme(legend.position = "bottom")
```


## Themes

`theme_bw` changes background to white instead of gray. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) + 
  labs(
    title = "Temperature plot",
    x = "Minimum daily temperature (C)",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) + 
  theme_bw() +
  theme(legend.position = "bottom")
```

## Learning Assessment

```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) + 
  geom_point(aes(size = prcp), alpha = .75) + 
  labs(
    title = "Temperature plot",
    x = "Date",
    y = "Maxiumum daily temperature (C)",
    caption = "Data from the rnoaa package"
  ) + 
  viridis::scale_color_viridis(discrete = TRUE) + 
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Arguments to `geom_*`:

Add in different dataset to the plot - stacking a new layer of new data: 

```{r}
central_park = 
  weather_df %>% 
  filter(name == "CentralPark_NY")

waikiki = 
  weather_df %>% 
  filter(name == "Waikiki_HA")

ggplot(waikiki, aes(x = date, y = tmax, color = name)) + 
  geom_point() + 
  geom_line(data = central_park)
```

