---
title: "Visualization Part 1"
author: "Devon Morgan"
date: "9/27/2018"
output: html_document
---

Code set up: 

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

# Set global parameters for figures produced today. 

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = 0.6,
  out.width = "90%"
)
```

## R Markdown


```{r data_import, cache = TRUE}
library(rnoaa)

weather_df = rnoaa::meteo_pull_monitors(c("USW00094728", "USC00519397", "USS0023B17S"),
                      var = c("PRCP", "TMIN", "TMAX"), 
                      date_min = "2017-01-01",
                      date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY", 
                      USC00519397 = "Waikiki_HA",
                      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
weather_df
```

## Revisit plot

Always helpful to include labels in plots. This is helpful for x and y axes and overall. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
      title = "Temperature plot",
      x = "Minimum daily temperature (C)",
      y = "Maxiumum daily temperature (C)",
      caption = "Data from the rnoaa package"
    )
```

Next, manipulate scale using `scale` functions. The `scale_x_*` where the value for * depends on the type of variable of interest. Use Option + 0 to make a degree symbol. 

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  labs(
      title = "Temperature plot",
      x = "Minimum daily temperature (C)",
      y = "Maxiumum daily temperature (C)",
      caption = "Data from the rnoaa package"
    ) + 
  scale_x_continuous(
      breaks = c(-15, 0, 15), 
      labels = c("-15ยบ C", "0", "15"))
```










```{r}
ggplot(weather_df, aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE) 


```

## Univariate Plots


```{r}
ggplot(weather_df, aes(x = tmax)) + 
  geom_histogram()


ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)

ggplot(weather_df, aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

Boxplots: 

```{r}
ggplot(weather_df, aes(x = name, y = tmax)) + geom_boxplot()

ggplot(weather_df, aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

## Last example: 

Make a bunch of plots: 

Boxplot does the best job of describing the distribution because can see more details about the data. 

```{r}
ggplot(weather_df, aes(y = prcp, x = name)) +
    geom_boxplot()
```

Violin plot is not as good because cannot see as much regarding the underlying data points. 

```{r}
ggplot(weather_df, aes(y = prcp, x = name)) +
    geom_violin()
```

```{r}
weather_df %>% 
    filter(prcp < 100) %>% 
ggplot(aes(x = prcp, fill = name)) + 
    geom_density(alpha = 0.5)
```

This version shows more of the data. 
```{r}
weather_df %>% 
    filter(prcp > 0) %>% 
ggplot(aes(x = prcp, fill = name)) + 
    geom_density(alpha = 0.5)
```



